Peaks and valleys of prolatin-related gene expression during pigeon parental care stages

### 2019-09-07 docs/characterization_manuscript.Rmd
looking better

##  Rock dove housing

@@

@@

EdgeR and DESeq2 were used to calculate statistical differnces in gene expression between each transition. All subsequent analyses were conducting using variance stabilized data from the DESeq2 model. A linear discriminant analysis was used to determine the confidence with which we could predict the parental care stage of a sample based on it's global gene expression profile. WGCNA was used to identify genes who's expression was correlated with specific candidate genes of interest.

Thosands of genes were found to be differentially expressed between the non-parental control group and the nest-building group (Table 1, Fig. 1). Within the time course of parental care, fewer than 1000 genes change there expression over the course of a few days, with the most pronounced difference in toltal number of changes occuring in the female pituitary between incubation days 9 and 17 (Table 1, Fig. 2).

@@

Even though we have a large sample size of roughly 12 samples per tissue per sex, the differences between all 8 sampled parental time points are so small, that we are only able to predict the stage using gene expression along 40%, 30% and 27% of the time in the pituiary, gonads and hypothalamus, respectively (Fig 3A). However, if we group the parental stages into anticipatory (control, nest building), egg care (lay and incubation days), and nestling care (hatch and nesting days), we incrase our ability to detect parental stage to 47%, 43%, and 37 pituiary, hypothalamus and gonads, respectively (Fig 3B). The

@@

Gene expression of prlactin (_PRL_)in the pitutiary, but not the hypothalamus follows the characteristic dip in expression relative to non-parental controls during early incubation then spikes the day before hatch and peaks at hatch (Fig 4). Intrestingly, hundreds of genes follow this same, prolactin-associated pattern of expression in the pitutiary (Fig. 5).

>>>>

@@

We used EdgeR determine the greatest contributor to variance in gene expression in the full dataset. Then, we used EdgeR and DESeq2 were used to calculate statistical differnces in gene expression between each transition within a given tissue. All subsequent analyses were conducting using variance stabilized data from the DESeq2 model. A linear discriminant analysis was used to determine the confidence with which we could predict the parental care stage of a sample based on it's global gene expression profile. WGCNA was used to identify genes who's expression was correlated with specific candidate genes of interest.

We found that the majority of the variation in the full dataset could be attributed to tissue, with only a notable sex differnce existing between the female ovaries and the male testies (Fig. 2).

Thousands of genes were found to be differentially expressed between the non-parental control group and the nest-building group (Table 1, Fig. 3). Within the time course of parental care, fewer than 1000 genes change there expression over the course of a few days, with the most pronounced difference in toltal number of changes occuring in the female pituitary between incubation days 9 and 17 (Table 1, Fig. 3).

Even though we have a large sample size of roughly 12 samples per tissue per sex, the differences between all 8 sampled parental time points are so small, that we are only able to predict the stage using gene expression along 40%, 30% and 27% of the time in the pituiary, gonads and hypothalamus, respectively (Fig 4A). However, if we group the parental stages into anticipatory (control, nest building), egg care (lay and incubation days), and nestling care (hatch and nesting days), we incrase our ability to detect parental stage to 47%, 43%, and 37 pituiary, hypothalamus and gonads, respectively (Fig 4B).

Gene expression of prlactin (_PRL_)in the pitutiary, but not the hypothalamus follows the characteristic dip in expression relative to non-parental controls during early incubation then spikes the day before hatch and peaks at hatch (Fig 5).

Hundreds of genes follow this same, prolactin-associated pattern of expression in the pitutiary (Fig. 6).

@@

```{r, fig.width= 5, echo = F, fig.cap = "MDS plots of all samples."}
ggdraw() + draw_image("../figures/limma/plotMDS-colored-1.png") 
```

@@

@@
